### 2014-06-05 src/s7_series.tex
Created placeholders

\chapter*{Donders series}

\begin{enumerate}
\item Van Aalderen-Smeets, S.I. (2007). \emph{Neural dynamics of visual selection.} Maastrict University, Maastricht, The Netherlands
\end{enumerate}