### 2014-06-05 src/s5_publications.tex
Updated style

\clearpage
\pagestyle{empty}

\chapter*{List of publications}
\phantomsection
\addcontentsline{toc}{chapter}{List of publications}

About me