### 2014-06-05 src/s7_series.tex

\clearpage
\pagestyle{empty}

\chapter*{Donders Graduate School for Cognitive Neuroscience Series}
\phantomsection
\addcontentsline{toc}{chapter}{Donders Series}%

%\noindent \begrip{}
\npar\npar
\footnotesize
\noindent 1.	Van Aalderen-Smeets, S.I. (2007). \emph{Neural dynamics of visual selection.} Maastricht University, Maastricht, the Netherlands.\npar \noindent 2.	Schoffelen, J.M. (2007). \emph{Neuronal communication through coherence in the human motor system.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 3.	De Lange, F.P. (2008). \emph{Neural mechanisms of motor imagery.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 4.	Grol, M.J. (2008). \emph{Parieto-frontal circuitry in visuomotor control.} Utrecht University, Utrecht, the Netherlands.\npar \noindent 5.	Bauer, M. (2008). \emph{Functional roles of rhythmic neuronal activity in the human visual and somatosensory system.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 6.	Mazaheri, A. (2008). \emph{The influence of ongoing oscillatory brain activity on evoked responses and behaviour.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 7.	Hooijmans, C.R. (2008). \emph{Impact of nutritional lipids and vascular factors in Alzheimer's disease.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 8.	Gaszner, B. (2008). \emph{Plastic responses to stress by the rodent urocortinergic Edinger-Westphal nucleus.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 9.	Willems, R.M. (2009). \emph{Neural reflections of meaning in gesture, language and action.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 10.	Van Pelt, S. (2009). \emph{Dynamic neural representations of human visuomotor space.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 11.	Lommertzen, J. (2009). \emph{Visuomotor coupling at different levels of complexity.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 12.	Poljac, E. (2009). \emph{Dynamics of cognitive control in task switching: Looking beyond the switch cost.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 13.	Poser, B.A. (2009). \emph{Techniques for BOLD and blood volume weighted fMRI.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 14.	Baggio, G. (2009). \emph{Semantics and the electrophysiology of meaning. Tense, aspect, event structure.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 15.	Van Wingen, G.A. (2009). \emph{Biological determinants of amygdala functioning.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 16.	Bakker, M. (2009). \emph{Supraspinal control of walking: Lessons from motor imagery.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 17.	Aarts, E. (2009). \emph{Resisting temptation: The role of the anterior cingulate cortex in adjusting cognitive control.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 18.	Prinz, S. (2009). \emph{Waterbath stunning of chickens – Effects of electrical parameters on the electroencephalogram and physical reflexes of broilers.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 19.	Knippenberg, J.M.J. (2009). \emph{The N150 of the Auditory Evoked Potential from the rat amygdala: In search for its functional significance.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 
20.	Dumont, G.J.H. (2009). \emph{Cognitive and physiological effects of 3,4-methylenedioxymethamphetamine (MDMA or 'ecstasy') in combination with alcohol or cannabis in humans.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 21.	Pijnacker, J. (2010). \emph{Defeasible inference in autism: A behavioral and electrophysiogical approach.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 22.	De Vrijer, M. (2010). \emph{Multisensory integration in spatial orientation.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 23.	Vergeer, M. (2010). \emph{Perceptual visibility and appearance: Effects of color and form.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 24.	Levy, J. (2010). \emph{In cerebro unveiling unconscious mechanisms during reading.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 25.	Treder, M. S. (2010). \emph{Symmetry in (inter)action.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 26.	Horlings C.G.C. (2010). \emph{A weak balance: Balance and falls in patients with neuromuscular disorders.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 27.	Snaphaan, L.J.A.E. (2010). \emph{Epidemiology of post-stroke behavioural consequences.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.  \npar \noindent 
28.	Dado-Van Beek, H.E.A. (2010). \emph{The regulation of cerebral perfusion in patients with Alzheimer's disease.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.  \npar \noindent 29.	Derks, N.M. (2010). \emph{The role of the non-preganglionic Edinger-Westphal nucleus in sex-dependent stress adaptation in rodents.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 30.	Wyczesany, M. (2010). \emph{Covariation of mood and brain activity. Integration of subjective self-report data with quantitative EEG measures.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 31.	Beurze S.M. (2010). \emph{Cortical mechanisms for reach planning.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 32.	Van Dijk, J.P. (2010). \emph{On the Number of Motor Units.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 33.	Lapatki, B.G. (2010). \emph{The Facial Musculature - Characterization at a Motor Unit Level.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 34.	Kok, P. (2010). \emph{Word order and verb inflection in agrammatic sentence production.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 35.	van Elk, M. (2010). \emph{Action semantics: Functional and neural dynamics.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 36.	Majdandzic, J. (2010). \emph{Cerebral mechanisms of processing action goals in self and others.} Radboud University Nijmegen, Nijmegen, the Netherlands.  \npar \noindent 37.	Snijders, T.M. (2010). \emph{More than words - Neural and genetic dynamics of syntactic unification.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 38.	Grootens, K.P. (2010). \emph{Cognitive dysfunction and effects of antipsychotics in schizophrenia and borderline personality disorder.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 39.	Nieuwenhuis, I.L.C. (2010). \emph{Memory consolidation: A process of integration – Converging evidence from MEG, fMRI and behavior.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 40.	Menenti, L.M.E. (2010). \emph{The right language: Differential hemispheric contributions to language production and comprehension in context.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 41.	Van Dijk, H.P. (2010). \emph{The state of the brain, how alpha oscillations shape behaviour and event related responses.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 42.	Meulenbroek, O.V. (2010). \emph{Neural correlates of episodic memory in healthy aging and Alzheimer's disease.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 43.	Oude Nijhuis, L.B. (2010). \emph{Modulation of human balance reactions.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 44.	Qin, S. (2010). \emph{Adaptive memory: Imaging medial temporal and prefrontal memory systems.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 45.	Timmer, N.M. (2011). \emph{The interaction of heparan sulfate proteoglycans with the amyloid protein.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 
46.	Craj\'e, C. (2011). \emph{(A)typical motor planning and motor imagery.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 47.	Van Grootel, T.J. (2011). \emph{On the role of eye and head position in spatial localisation behaviour.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 48.	Lamers, M.J.M. (2011). \emph{Levels of selective attention in action planning.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 49.	Van der Werf, J. (2011). \emph{Cortical oscillatory activity in human visuomotor integration.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 50.	Scheeringa, R. (2011). \emph{On the relation between oscillatory EEG activity and the BOLD signal.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 
51.	B\"ogels, S. (2011). \emph{The role of prosody in language comprehension: When prosodic breaks and pitch accents come into play.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 52.	Ossewaarde, L. (2011). \emph{The mood cycle: Hormonal influences on the female brain.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 53.	Kuribara, M. (2011). \emph{Environment-induced activation and growth of pituitary melanotrope cells of Xenopus laevis.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 54.	Helmich, R.C.G. (2011). \emph{Cerebral reorganization in Parkinson's disease.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 55.	Boelen, D. (2011). \emph{Order out of chaos? Assessment and treatment of executive disorders in brain-injured patients.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 56.	Koopmans, P.J. (2011). \emph{fMRI of cortical layers.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 57.	van der Linden, M.H. (2011). \emph{Experience-based cortical plasticity in object category representation.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 58.	Kleine, B.U. (2011). \emph{Motor unit discharges - Physiological and diagnostic studies in ALS.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands. \npar \noindent 59.	Paulus, M. (2011). \emph{Development of action perception: Neurocognitive mechanisms underlying children's processing of others' actions.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 60.	Tieleman, A.A. (2011). \emph{Myotonic dystrophy type 2. A newly diagnosed disease in the Netherlands.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands. \npar \noindent 61.	Van Leeuwen, T.M. (2011). \emph{'How one can see what is not there': Neural mechanisms of grapheme-colour synaesthesia.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 62.	Van Tilborg, I.A.D.A. (2011). \emph{Procedural learning in cognitively impaired patients and its application in clinical practice.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 63.	Bruinsma, I.B. (2011). \emph{Amyloidogenic proteins in Alzheimer's disease and Parkinson's disease: Interaction with chaperones and inflammation.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 64.	Voermans, N. (2011). \emph{Neuromuscular features of Ehlers-Danlos syndrome and Marfan syndrome; expanding the phenotype of inherited connective tissue disorders and investigating the role of the extracellular matrix in muscle.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands. \npar \noindent 65.	Reelick, M. (2011). \emph{One step at a time. Disentangling the complexity of preventing falls in frail older persons.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands. \npar \noindent 66.	Buur, P.F. (2011). \emph{Imaging in motion. Applications of multi-echo fMRI.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 67.	Schaefer, R.S. (2011). \emph{Measuring the mind's ear: EEG of music imagery.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 68.	Xu, L. (2011). \emph{The non-preganglionic Edinger-Westphal nucleus: An integration center for energy balance and stress adaptation.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 69.	Schellekens, A.F.A.  (2011). \emph{Gene-environment interaction and intermediate phenotypes in alcohol dependence.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 70.	Van Marle, H.J.F. (2011). \emph{The amygdala on alert: A neuroimaging investigation into amygdala function during acute stress and its aftermath.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 71.	De Laat, K.F. (2011). \emph{Motor performance in individuals with cerebral small vessel disease: An MRI study.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.
\npar \noindent 72.	M\"adebach, A. (2011). \emph{Lexical access in speaking: Studies on lexical selection and cascading activation.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 73.	Poelmans, G.J.V. (2011). \emph{Genes and protein networks for neurodevelopmental disorders.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 74.	Van Norden, A.G.W. (2011). \emph{Cognitive function in elderly individuals with cerebral small vessel disease. An MRI study.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 75.	Jansen, E.J.R. (2011). \emph{New insights into V-ATPase functioning: the role of its accessory subunit Ac45 and a novel brain-specific Ac45 paralog.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 76.	Haaxma, C.A. (2011). \emph{New perspectives on preclinical and early stage Parkinson's disease.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 77.	Haegens, S. (2012). \emph{On the functional role of oscillatory neuronal activity in the somatosensory system.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 78.	van Barneveld, D.C.P.B.M. (2012). \emph{Integration of exteroceptive and interoceptive cues in spatial localization.} Radboud University Nijmegen, Nijmegen, the Netherlands. \npar \noindent 79.	Spies, P.E. (2012). \emph{The reflection of Alzheimer disease in CSF.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 80.	Helle, M. (2012). \emph{Artery-specific perfusion measurements in the cerebral vasculature by magnetic resonance imaging.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 81.	Egetemeir, J. (2012). \emph{Neural correlates of real-life joint action.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 82.	Janssen, L. (2012). \emph{Planning and execution of (bi)manual grasping.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 83.	Vermeer, S. (2012). \emph{Clinical and genetic characterisation of autosomal recessive cerebellar ataxias.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 84.	Vrins, S. (2012). \emph{Shaping object boundaries: Contextual effects in infants and adults.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 85.	Weber, K.M. (2012). \emph{The language learning brain: Evidence from second language and bilingual studies of syntactic processing.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 86.	Verhagen, L. (2012). \emph{How to grasp a ripe tomato.} Utrecht University, Utrecht, the Netherlands.\npar \noindent 87.	Nonkes, L.J.P. (2012). \emph{Serotonin transporter gene variance causes individual differences in rat behaviour: For better and for worse.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 88.	Joosten-Weyn Banningh, L.W.A. (2012). \emph{Learning to live with Mild Cognitive Impairment: development and evaluation of a psychological intervention for patients with Mild Cognitive Impairment and their significant others.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 89.	Xiang, HD. (2012). \emph{The language networks of the brain.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 90.	Snijders, A.H. (2012). \emph{Tackling freezing of gait in Parkinson's disease.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 91.	Rouwette, T.P.H. (2012). \emph{Neuropathic pain and the brain - Differential involvement of corticotropin-releasing factor and urocortin 1 in acute and chronic pain processing.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 92.	Van de Meerendonk, N. (2012). \emph{States of indecision in the brain: Electrophysiological and hemodynamic reflections of monitoring in visual language perception.} Radboud University Nijmegen, Nijmegen, the Netherlands.\npar \noindent 93.	Sterrenburg, A. (2012). \emph{The stress response of forebrain and midbrain regions: Neuropeptides, sex-specificity and epigenetics.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 94.	Uithol, S. (2012). \emph{Representing action and intention.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 95.	Van Dam, W.O.  (2012). \emph{On the specificity and flexibility of embodied lexical-semantic representations.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 96.	Slats, D. (2012). \emph{ CSF biomarkers of Alzheimer's disease: Serial sampling analysis and the study of circadian rhythmicity.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 97.	Van Nuenen, B.F.L. (2012). \emph{Cerebral reorganization in premotor parkinsonism.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 98.	van Schouwenburg, M.R. (2012). \emph{Fronto-striatal mechanisms of attentional control.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 99.	Azar, M.G. (2012). \emph{On the theory of reinforcement learning: Methods, convergence analysis and sample complexity.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 100.	Meeuwissen, E.B. (2012). \emph{Cortical oscillatory activity during memory formation.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 101.	Arnold, J.F. (2012). \emph{When mood meets memory: Neural and behavioral perspectives on emotional memory in health and depression.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 102.	Gons, R.A.R. (2012). \emph{Vascular risk factors in cerebral small vessel disease: A diffusion tensor imaging study.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 103.	Wingbermühle, E. (2012). \emph{Cognition and emotion in adults with Noonan syndrome: A neuropsychological perspective.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 104.	Walentowska, W. (2012). \emph{Facing emotional faces. The nature of automaticity of facial emotion processing studied with ERPs.} Radboud University Nijmegen, Nijmegen, The Netherlands. \npar \noindent 105.	Hoogman, M. (2012). \emph{Imaging the effects of ADHD risk genes.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 106.	Tramper, J. J. (2012). \emph{Feedforward and feedback mechanisms in sensory motor control.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 107.	Van Eijndhoven, P. (2012). \emph{State and trait characteristics of early course major depressive disorder.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands.\npar \noindent 108.	Visser, E. (2012). \emph{Leaves and forests: Low level sound processing and methods for the large-scale analysis of white matter structure in autism.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 109.	Van Tooren-Hoogenboom, N. (2012). \emph{Neuronal communication in the synchronized brain. Investigating the functional role of visually-induced gamma band activity: Lessons from MEG.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 110.	Henckens, M.J.A.G. (2012). \emph{Imaging the stressed brain. Elucidating the time- and region-specific effects of stress hormones on brain function: A translational approach.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 111.	Van Kesteren, M.T.R. (2012). \emph{Schemas in the brain: Influences of prior knowledge on learning, memory, and education.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 112.	Brenders, P. (2012). \emph{Cross-language interactions in beginning second language learners.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 113.	Ter Horst, A.C. (2012). \emph{Modulating motor imagery. Contextual, spatial and kinaesthetic influences.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 114.	Tesink, C.M.J.Y. (2013). \emph{Neurobiological insights into language comprehension in autism: Context matters.} Radboud University Nijmegen, Nijmegen, The Netherlands.
\npar \noindent 115.	B\"ockler, A. (2013). \emph{Looking at the world together. How others' attentional relations to jointly attended scenes shape cognitive processing.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 116.	Van Dongen, E.V. (2013). \emph{Sleeping to Remember. On the neural and behavioral mechanisms of sleep-dependent memory consolidation.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 117.	Volman, I. (2013). \emph{The neural and endocrine regulation of emotional actions.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 118.	Buchholz, V. (2013). \emph{Oscillatory activity in tactile remapping.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 119.	Van Deurzen, P.A.M. (2013). \emph{Information processing and depressive symptoms in healthy adolescents.} Radboud University Nijmegen, Nijmegen, The Netherlands.
\npar \noindent 120.	Whitmarsh, S. (2013). \emph{Nonreactivity and metacognition in mindfulness.} Radboud University Nijmegen, Nijmegen, The Netherlands.
\npar \noindent 121.	Vesper, C. (2013). \emph{Acting together: Mechanisms of intentional coordination.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 122.	Lagro, J. (2013). \emph{Cardiovascular and cerebrovascular physiological measurements in clinical practice and prognostics in geriatric patients.} Radboud University Nijmegen Medical Centre, Nijmegen, the Netherlands. \npar \noindent 123.	Eskenazi, T.T. (2013). \emph{ You, us \& them: From motor simulation to ascribed shared intentionality in social perception.} Radboud University Nijmegen, Nijmegen, The Netherlands. \npar \noindent 124.	Ondobaka, S. (2013). \emph{On the conceptual and perceptual processing of own and others' behavior.} Radboud University Nijmegen, Nijmegen, The Netherlands. \npar \noindent 125.	Overvelde, J.A.A.M. (2013). \emph{Which practice makes perfect? Experimental studies on the acquisition of movement sequences to identify the best learning condition in good and poor writers.} Radboud University Nijmegen, Nijmegen, The Netherlands. \npar \noindent 126.	Kalisvaart, J.P. (2013). \emph{Visual ambiguity in perception and action.} Radboud University Nijmegen Medical Centre, Nijmegen, The Netherlands. \npar \noindent 127.	Kroes, M. (2013). \emph{Altering memories for emotional experiences.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 128.	Duijnhouwer, J. (2013). \emph{ Studies on the rotation problem in self-motion perception.} Radboud University Nijmegen, Nijmegen, The Netherlands.
\npar \noindent 129.	Nijhuis, E.H.J (2013). \emph{ Macroscopic networks in the human brain: Mapping connectivity in healthy and damaged brains.} University of Twente, Enschede, The Netherlands.
\npar \noindent 130.	Braakman, M. H. (2013). \emph{Posttraumatic stress disorder with secondary psychotic features. A diagnostic validity study among refugees in the Netherlands.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 131.	Zedlitz, A.M.E.E. (2013). \emph{Brittle brain power. Post-stroke fatigue, explorations into assessment and treatment.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 132.	Schoon, Y. (2013). \emph{From a gait and falls clinic visit towards self-management of falls in frail elderly.} Radboud University Nijmegen Medical Centre, Nijmegen, The Netherlands.\npar \noindent 133.	Jansen, D. (2013). \emph{The role of nutrition in Alzheimer's disease - A study in transgenic mouse models for Alzheimer's disease and vascular disorders.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 134.	Kos, M. (2013). \emph{On the waves of language - Electrophysiological reflections on semantic and syntactic processing.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 135.	Severens, M. (2013). \emph{Towards clinical BCI applications: Assistive technology and gait rehabilitation.} Radboud University Nijmegen, Nijmegen, Sint Maartenskliniek, Nijmegen, The Netherlands.\npar \noindent 136.	Bergmann, H. (2014). \emph{Two is not always better than one: On the functional and neural (in)dependence of working memory and long-term memory.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 137.	Wronka, E. (2013). \emph{Searching for the biological basis of human mental abilitites. The relationship between attention and intelligence studied with P3.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 138.	L\"uttjohann, A.K. (2013). \emph{The role of the cortico-thalamo-cortical system in absence epilepsy.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 139.	Brazil, I.A. (2013). \emph{Change doesn't come easy: Dynamics of adaptive behavior in psychopathy.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 140.	Zerbi, V. (2013). \emph{ Impact of nutrition on brain structure and function. A magnetic resonance imaging approach in Alzheimer mouse models.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 141.	Delnooz, C.C.S. (2014). \emph{Unravelling primary focal dystonia. A treatment update and new pathophysiological insights.} Radboud University Nijmegen Medical Centre, Nijmegen, The Netherlands.\npar \noindent 142.	Bultena, S.S. (2013). \emph{Bilingual processing of cognates and language switches in sentence context.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 143.	Janssen, G. (2014). \emph{Diagnostic assessment of psychiatric patients: A contextual perspective on executive functioning.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent 144.	Piai, V. Magalh\~aes (2014). \emph{Choosing our words: Lexical competition and the involvement of attention in spoken word production.} Radboud University Nijmegen, Nijmegen, The Netherlands.\npar \noindent \npar\npar \noindent \npar