### 2014-09-09 src/s5_publications.tex

\clearpage
\pagestyle{empty}

\chapter*{List of publications}
\phantomsection
\addcontentsline{toc}{chapter}{List of publications}

\begin{enumerate}

\item \textbf{Clemens, I. A. H.}, De Vrijer, M., Van Gisbergen, J. A. M. \& Medendorp, W. P.
\citeyear{clemens2011}. Multisensory processing in spatial orientation: An inverse probabilistic approach. \textit{Journal of Neuroscience}, \textit{31}(14), 5365-5377.

\item \textbf{Clemens, I. A. H.}, Selen, L. P. J., Koppen, M. \& Medendorp, W. P. \citeyear{clemens2012}. Visual stability across combined eye and body motion. \textit{Journal of Vision}, \textit{12}(12).

\item Lodder, G. M. A., Scholte, R. H. J., \textbf{Clemens, I. A. H.}, Engels, R. C. M. E., Goossens, L \& Verhagen, M. \citeyear{lodder2015}. Loneliness and overt visual attention to social cues: An eye-tracking study. \textit{Submitted}

\item \textbf{Clemens, I. A. H.}, Selen, L. P. J., MacNeilage, P. R. \& Medendorp, W. P. \citeyear{clemens2015a}. Title. \textit{In preparation}

\item \textbf{Clemens, I. A. H.}, Selen, L. P. J., MacNeilage, P. R. \& Medendorp, W. P. \citeyear{clemens2015b}. Title. \textit{In preparation}

\end{enumerate}