### 2014-09-09 src/misc/s1_intro.tex
Moved files

\chapter{General introduction}
\newpage

The vestibular system is composed of two parts: the otoliths and the semicircular canals.

The otoliths are sensitive to linear accelearion with can in turn be cause by gravity and interial forces.

These are therefore important for the perception of: verticality, heading and displacement.

The canals are sensitive for rotational velocity and are important for

Discuss TVOR / RVOR and relation with otoliths / canals. Neural mechanism?

Paper 1 - Linking SVV (space) and SBT (body-in-space)
Paper 3 - Influence of eye movements on self-motion perception
Paper 4 - Influence of fixation depth on self-motion perception
Paper 2 - Updating of visual stimuli across lateral translation

Overlapping themes:
  Vestibular, only otoliths; gravity (paper 1), linear acceleration (paper 2 to 4)
  Visual, orientation judgment (paper 1), updating (paper 2), self-motion (paper 3, 4)