### 2014-09-09 src/misc/s3_summary.tex

\clearpage
\pagestyle{empty}

\chapter*{Nederlandse samenvatting}
\phantomsection
\addcontentsline{toc}{chapter}{Nederlandse samenvatting}

Summary