### 2014-09-09 src/paper1.tex
Allow creating separate files for papers.

\documentclass[10pt,b5paper,twoside]{book}

\input{src/common.tex}

\hypersetup{
    pdfauthor = {Ivar Clemens, Maaike de Vrijer, Luc Selen, Jan van Gisbergen, and Pieter Medendorp},
    pdftitle = {Multisensory processing in spatial orientation:\\An inverse probabilistic approach},
    %pdfsubject = {Language production},
    %pdfkeywords = {language production; }
}

\begin{document}

\selectlanguage{british}
\input{src/paper1/paper1.tex}

\bibliography{refs}{}

\titleformat{\chapter}
{\normalfont\Large\bfseries}{\thechapter}{0.2em}{}
\titlespacing*{\chapter}{0pt}{-30pt}{10pt}

\end{document}