### 2014-09-09 src/paper2.tex

\documentclass[10pt,b5paper,twoside]{book}

\input{src/common.tex}

\hypersetup{
    pdfauthor = {Ivar Clemens, Luc Selen, Mathieu Koppen, and Pieter Medendorp},
    pdftitle = {Visual stability across combined eye and body motion}
}

\begin{document}

\selectlanguage{british}
\input{src/paper2/paper2.tex}

\bibliography{refs}{}

\titleformat{\chapter}
{\normalfont\Large\bfseries}{\thechapter}{0.2em}{}
\titlespacing*{\chapter}{0pt}{-30pt}{10pt}

\end{document}