### 2014-10-07 src/misc/s4_aboutme.tex
About me paragraph

\clearpage
\pagestyle{empty}

\chapter*{Curriculum vitae}
\phantomsection
\addcontentsline{toc}{chapter}{Curriculum vitae}

Ivar Clemens was born on the 2nd of June 1984 in Boxtel, The Netherlands. He started his studies in 2000 at the Koning Willem 1 College where he obtained a degree in Industrial Automation. He then started work on a master's degree in Biology at the Radboud University Nijmegen. As part of this master's degree Ivar did his research internship on the \emph{integration of pre- and post-vergence target distance cues} with Pieter Medendorp at the Nijmegen Institute for Cognition and Information (NICI, later renamed to the Donders Centre for Cognition). Ivar then switched to the Donders Centre for Cognitive Neuro-imaging (DCCN) for another research internship. This time he worked with Ole Jensen and Saskia Haegens on \emph{homeostatic plasticity in the somatosensory cortex}. After completing his master's degree in 2009, Ivar started his PhD project at he Donders Centre for Cognition. The results of this PhD project are contained within this thesis. As of [month] 2015, Ivar is employed at [institute] where he works on [topic].