### 2014-10-07 src/misc/s6_thanks.tex
Added test, updated style.

\clearpage
\pagestyle{empty}

\chapter*{Acknowledgments}
\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgments}