### 2014-10-07 src/test/figref.tex

\documentclass[10pt,b5paper,twoside]{book}

\input{src/common.tex}

\usepackage{src/my_thesis}

\begin{document}
\selectlanguage{british}

\chapter{Table references test}

\begin{itemize}
\item \commonref{ and }{Type}{p1:tab1}{B}
\item \figref{p1:tab1}
\item \figref{p1:tab2}
\item \tabref{p1:tab2}
\item \figref{p1:tab2}
\item \figref[ or ]{p1:tab1,p1:tab2}
\end{itemize}

\begin{table}

\begin{tabular}{lllllll}
\hline
Subject & $a_{HS}$ (\textdegree/\textdegree) & $b_{HS}$ (\textdegree) & $\sigma_{HSP}$ (\textdegree) & $\sigma_{BS}$ (\textdegree) & $\sigma_{HB}$ (\textdegree) & $A_{OCR}$ (\textdegree) \\
\hline
S1 & 0.23 \textpm 0.02 & 1.2 \textpm 0.32 & 11.6 \textpm 1.0 & 12.3 \textpm 1.1 & 3.3 \textpm 1.2 & 27.0 \textpm 2.2 \\
S2 & 0.12 \textpm 0.02 & 1.2 \textpm 0.52 & 9.4 \textpm 1.1 & 8.4 \textpm 2.9 & 6.4 \textpm 4.1 & 17.0 \textpm 3.8 \\
S3 & 0.20 \textpm 0.03 & 1.1 \textpm 0.42 & 14.4 \textpm 1.7 & 6.7 \textpm 1.9 & 9.3 \textpm 2.4 & 17.5 \textpm 2.1 \\
S4 & 0.07 \textpm 0.50 & 3.9 \textpm n/a & 11.2 \textpm 1.3 & 12.6 \textpm 2.3 & 7.1 \textpm 3.5 & 0 \textpm n/a \\
S5 & 0.11 \textpm n/a & 3.3 \textpm 1.0 & 18.7 \textpm 4.8 & 15.0 \textpm n/a & 3.6 \textpm 2.1 & 1.06 \textpm n/a \\
S6 & 0.23 \textpm 0.09 & 3.0 \textpm 1.5 & 9.5 \textpm 1.1 & 8.0 \textpm 0.83 & 1.8 \textpm n/a & 18.8 \textpm 4.1 \\
S7 & 0.20 \textpm 0.14 & 3.2 \textpm 1.0 & 12.8 \textpm 2.4 & 12.7 \textpm 6.1 & 3.0 \textpm n/a & 20.8 \textpm 9.0 \\
Mean & 0.16 \textpm 0.06 & 2.4 \textpm 1.2 & 12.5 \textpm 3.2 & 10.8 \textpm 3.1 & 4.9 \textpm 2.7 & 14.6 \textpm 10.2 \\
\end{tabular}

\caption{Best-fit parameter and bootstrap-based SD values. Imposed fit limits were as follows: $a_{HS}$: 0.5\textdegree/\textdegree; $b_{HS}$, $\sigma_{HSP}$, $\sigma_{BS}$, $\sigma_{HB}$, 50\textdegree; $A_{OCR}$, 30\textdegree. SD values are not shown (n/a) when bootstrapped values formed a skewed distribution. $a_{HS}$, Tilt-related increase in otolith noise; $b_{HS}$, otolith noise in upright position; $\sigma{HSP}$, width of head-tilt prior; $\sigma{BS}$, noise in body-in-space sensors; $\sigma{HB}$, noise in neck sensors; $A_{OCR}$, uncompensated ocular counterroll.}
\label{p1:tab1}
\end{table}

\begin{table}
\begin{tabular}{llllll}
\hline
Subject & MLE full model & MLE reduced model & $p$ & BIC full model & BIC psychometric fits \\
\hline
S1 & 231.5 & 312.4 & \textless 0.001 & 498.1 & 540.3 \\
S2 & 197.8 & 216.9 & \textless 0.001 & 430.6 & 539.3 \\
S3 & 267.5 & 313.4 & \textless 0.001 & 570.0 & 548.2 \\
S4 & 207.5 & 218.2 & \textless 0.001 & 450.0 & 569.1 \\
S5 & 195.6 & 216.3 & \textless 0.001 & 426.2 & 571.0 \\
S6 & 163.5 & 183.2 & \textless 0.001 & 362.0 & 502.5 \\
S7 & 248.0 & 260.0 & \textless 0.001 & 531.1 & 633.5 \\ 
\end{tabular}
\caption{Validation of the model. Log-likelihood ratio test of the full model (with indirect pathways) and reduced model (without indirect pathways). For all subjects, the full model outperforms the reduced model lacking multisensory integration through the indirect pathways. BIC values are much lower for the Bayesian integration model compared with separate psychometric fits in six of seven subjects.}
\label{p1:tab2}
\end{table}

\end{document}