### 2015-01-16 src/paper3/paper3.tex

\thispagestyle{empty}

\chapter{Integration of ocular and vestibular signals for self-motion perception in darkness}
\chaptermark{}

\newpage

\small {\bf Abstract} Self-motion is typically accompanied by compensatory eye movements that help minimize retinal slip and maximize dynamic visual acuity. To date, it is unknown whether these eye movements also have a reversed role, serving as a cue for self-motion perception. To address this question, we had participants ($n=8$) judge self-motion during different eye movement conditions in the absence of full-field optic flow.  In a 2-AFC task, participants indicated whether the second of two successive, passive, lateral whole-body translations was longer or shorter than the first. Eye movements during each translation were world-stationary, body-stationary or free in an otherwise dark room. Results show that the perceived translations were shorter with body-fixed gaze compared to world-fixed gaze. Using a linear model, we estimated the vestibular and eye movement contributions to self-motion perception and found that eye movement signals contribute approximately 25 percent. The model was validated independently by predicting the effects of eye movements on self-motion perception during the free eye movement conditions, taking the inter-subject variation in oculomotor weight and eye movement patterns into account. We conclude that eye movement signals influence self-motion perception, even in the absence of visual stimulation, and even when oculomotor and vestibular estimates are in conflict, e.g. during body-fixed gaze. We hypothesize that adverse consequences of this seemingly inflexible arrangement are minimal under natural conditions because eye movements and self-motion are highly correlated, and because eye movements are most often accompanied by veridical optic flow cues to self-motion.

\vfill

\noindent\underline{ \hspace{4cm} }

\noindent This chapter has been published as \newline
\noindent {\bf Clemens, I.A.H.}, Selen, L.P.J., MacNeilage, P.R. and Medendorp, W.P. \citeyear{clemens2015a}. Title. \emph{Journal}, volume(issue): from-to. \newline

\newpage

\input{src/paper3/p3_intro}
\input{src/paper3/p3_methods}
\input{src/paper3/p3_results}
\input{src/paper3/p3_discussion}