### 2015-01-16 src/paper4/paper4.tex

\thispagestyle{empty}

% Incomplete compensation for fixation depth in eye movement driven self-motion perception

\chapter{Incomplete compensation for fixation depth for eye movement induced self-motion perception}
\chaptermark{}

\newpage

\small {\bf Abstract}

\vfill

\noindent\underline{ \hspace{4cm} }

%\noindent This chapter has been published as \newline
%\noindent {\bf Clemens, I.A.H.}, Selen, L.P.J., MacNeilage, P.R. and Medendorp, W.P. \citeyear{clemens2015b}. Title. \emph{Journal}, volume(issue): from-to. \newline

\newpage

\input{src/paper4/p4_intro}
\input{src/paper4/p4_methods}
\input{src/paper4/p4_results}
\input{src/paper4/p4_discussion}