### 2015-03-21 src/paper4/paper4.tex

\thispagestyle{empty}

% Incomplete compensation for fixation depth in eye movement driven self-motion perception

\chapter{Incomplete compensation for fixation depth for eye movement induced self-motion perception}
\chaptermark{}

\newpage

\small {\bf Abstract}
Previously, we have shown that the eye movements that typically accompany self-motion are used to augment self-motion perception. As fixation depth modulates the amplitude of these eye movements, it should be taken into account for a veridical oculomotor estimate of self-motion. To investigate whether the brain compensates for fixation depth, we had  participants ($n = 8$) judge self-motion during different eye movement conditions in the absence of full-field optic flow.  In a 2-AFC task, participants indicated whether the second of two successive passive lateral whole-body translations was longer or shorter than the first. At the start of each movement participants fixated either a nearby or a far away target, which was either body stationary or world stationary during the translation interval. Results show that the perceived translations were shorter for nearby world-fixed gaze compared to far away world-fixed gaze, indicating that eye movements are not properly scaled in self-motion perception. Together with the observation that self-motion perception is not affected by the depth of a body stationary fixation target, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

\vfill

\noindent\underline{ \hspace{4cm} }

\noindent This chapter is being prepared for publication \newline
\noindent {\bf Clemens, I.A.H.}, Selen, L.P.J., MacNeilage, P.R. and Medendorp, W.P. \citeyear{clemens2015b}. %Title. \emph{Journal}, volume(issue): from-to. \newline

\newpage

\input{src/paper4/p4_intro}
\input{src/paper4/p4_methods}
\input{src/paper4/p4_results}
\input{src/paper4/p4_discussion}