### 2015-07-08 src/misc/s7_series_list.tex
Generate Donders series tex.

\dientry{1}{Van Aalderen-Smeets, S.I.}{2007}
	{Neural dynamics of visual selection.}
	{Maastricht University}{Maastricht}{the Netherlands}{}
\dientry{2}{Schoffelen, J.M.}{2007}
	{Neuronal communication through coherence in the human motor system.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/55760}
\dientry{3}{De Lange, F.P.}{2008}
	{Neural mechanisms of motor imagery.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/73602}
\dientry{4}{Grol, M.J.}{2008}
	{Parieto-frontal circuitry in visuomotor control.}
	{Utrecht University}{Utrecht}{the Netherlands}{}
\dientry{5}{Bauer, M.}{2008}
	{Functional roles of rhythmic neuronal activity in the human visual and somatosensory system.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/72591}
\dientry{6}{Mazaheri, A.}{2008}
	{The influence of ongoing oscillatory brain activity on evoked responses and behaviour.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{7}{Hooijmans, C.R.}{2008}
	{Impact of nutritional lipids and vascular factors in Alzheimer’s disease.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{8}{Gaszner, B.}{2008}
	{Plastic responses to stress by the rodent urocortinergic Edinger-Westphal nucleus.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{9}{Willems, R.M.}{2009}
	{Neural reflections of meaning in gesture, language and action.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{10}{Van Pelt, S.}{2009}
	{Dynamic neural representations of human visuomotor space.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{11}{Lommertzen, J.}{2009}
	{Visuomotor coupling at different levels of complexity.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{12}{Poljac, E.}{2009}
	{Dynamics of cognitive control in task switching: Looking beyond the switch cost.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{13}{Poser, B.A.}{2009}
	{Techniques for BOLD and blood volume weighted fMRI.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{14}{Baggio, G.}{2009}
	{Semantics and the electrophysiology of meaning. Tense, aspect, event structure.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{15}{Van Wingen, G.A.}{2009}
	{Biological determinants of amygdala functioning.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{16}{Bakker, M.}{2009}
	{Supraspinal control of walking: Lessons from motor imagery.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{2066/74406}
\dientry{17}{Aarts, E.}{2009}
	{Resisting temptation: The role of the anterior cingulate cortex in adjusting cognitive control.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{18}{Prinz, S.}{2009}
	{Waterbath stunning of chickens – Effects of electrical parameters on the electroencephalogram and physical reflexes of broilers.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{19}{Knippenberg, J.M.J.}{2009}
	{The N150 of the Auditory Evoked Potential from the rat amygdala: In search for its functional significance.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{20}{Dumont, G.J.H.}{2009}
	{Cognitive and physiological effects of 3,4-methylenedioxymethamphetamine (MDMA or ’ecstasy’) in combination with alcohol or cannabis in humans.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{21}{Pijnacker, J.}{2010}
	{Defeasible inference in autism: A behavioral and electrophysiogical approach.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/76509}
\dientry{22}{De Vrijer, M.}{2010}
	{Multisensory integration in spatial orientation.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{23}{Vergeer, M.}{2010}
	{Perceptual visibility and appearance: Effects of color and form.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{24}{Levy, J.}{2010}
	{In cerebro unveiling unconscious mechanisms during reading.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/76500}
\dientry{25}{Treder, M. S.}{2010}
	{Symmetry in (inter)action.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{26}{Horlings C.G.C.}{2010}
	{A weak balance: Balance and falls in patients with neuromuscular disorders.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{27}{Snaphaan, L.J.A.E.}{2010}
	{Epidemiology of post-stroke behavioural consequences.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{28}{Dado – Van Beek, H.E.A.}{2010}
	{The regulation of cerebral perfusion in patients with Alzheimer’s disease.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{29}{Derks, N.M.}{2010}
	{The role of the non-preganglionic Edinger-Westphal nucleus in sex-dependent stress adaptation in rodents.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{30}{Wyczesany, M.}{2010}
	{Covariation of mood and brain activity. Integration of subjective self-report data with quantitative EEG measures.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{31}{Beurze S.M.}{2010}
	{Cortical mechanisms for reach planning.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/77571}
\dientry{32}{Van Dijk, J.P.}{2010}
	{On the Number of Motor Units.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{33}{Lapatki, B.G.}{2010}
	{The Facial Musculature - Characterization at a Motor Unit Level.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{34}{Kok, P.}{2010}
	{Word order and verb inflection in agrammatic sentence production.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{35}{van Elk, M.}{2010}
	{Action semantics: Functional and neural dynamics.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{36}{Majdandzic, J.}{2010}
	{Cerebral mechanisms of processing action goals in self and others.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/77568}
\dientry{37}{Snijders, T.M.}{2010}
	{More than words - Neural and genetic dynamics of syntactic unification.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{38}{Grootens, K.P.}{2010}
	{Cognitive dysfunction and effects of antipsychotics in schizophrenia and borderline personality disorder.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{39}{Nieuwenhuis, I.L.C.}{2010}
	{Memory consolidation: A process of integration – Converging evidence from MEG, fMRI and behavior.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{40}{Menenti, L.M.E.}{2010}
	{The right language: Differential hemispheric contributions to language production and comprehension in context.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{41}{Van Dijk, H.P.}{2010}
	{The state of the brain, how alpha oscillations shape behaviour and event related responses.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/81998}
\dientry{42}{Meulenbroek, O.V.}{2010}
	{Neural correlates of episodic memory in healthy aging and Alzheimer’s disease.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/83219}
\dientry{43}{Oude Nijhuis, L.B.}{2010}
	{Modulation of human balance reactions.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/83249}
\dientry{44}{Qin, S.}{2010}
	{Adaptive memory: Imaging medial temporal and prefrontal memory systems.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{45}{Timmer, N.M.}{2011}
	{The interaction of heparan sulfate proteoglycans with the amyloid protein.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{46}{Crajé, C.}{2011}
	{(A)typical motor planning and motor imagery.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{47}{Van Grootel, T.J.}{2011}
	{On the role of eye and head position in spatial localisation behaviour.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/83262}
\dientry{48}{Lamers, M.J.M.}{2011}
	{Levels of selective attention in action planning.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{49}{Van der Werf, J.}{2011}
	{Cortical oscillatory activity in human visuomotor integration.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/85057}
\dientry{50}{Scheeringa, R.}{2011}
	{On the relation between oscillatory EEG activity and the BOLD signal.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{51}{Bögels, S.}{2011}
	{The role of prosody in language comprehension: When prosodic breaks and pitch accents come into play.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/83279}
\dientry{52}{Ossewaarde, L.}{2011}
	{The mood cycle: Hormonal influences on the female brain.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{53}{Kuribara, M.}{2011}
	{Environment-induced activation and growth of pituitary melanotrope cells of Xenopus laevis.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{54}{Helmich, R.C.G.}{2011}
	{Cerebral reorganization in Parkinson’s disease.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/87164}
\dientry{55}{Boelen, D.}{2011}
	{Order out of chaos? Assessment and treatment of executive disorders in brain-injured patients.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/87197}
\dientry{56}{Koopmans, P.J.}{2011}
	{fMRI of cortical layers.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{57}{van der Linden, M.H.}{2011}
	{Experience-based cortical plasticity in object category representation.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/79490}
\dientry{58}{Kleine, B.U.}{2011}
	{Motor unit discharges - Physiological and diagnostic studies in ALS.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{59}{Paulus, M.}{2011}
	{Development of action perception: Neurocognitive mechanisms underlying children’s processing of others’ actions.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/87194}
\dientry{60}{Tieleman, A.A.}{2011}
	{Myotonic dystrophy type 2. A newly diagnosed disease in the Netherlands.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{61}{Van Leeuwen, T.M.}{2011}
	{‘How one can see what is not there’: Neural mechanisms of grapheme-colour synaesthesia.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/74392}
\dientry{62}{Van Tilborg, I.A.D.A.}{2011}
	{Procedural learning in cognitively impaired patients and its application in clinical practice.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{63}{Bruinsma, I.B.}{2011}
	{Amyloidogenic proteins in Alzheimer’s disease and Parkinson’s disease: Interaction with chaperones and inflammation.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{64}{Voermans, N.}{2011}
	{Neuromuscular features of Ehlers-Danlos syndrome and Marfan syndrome; expanding the phenotype of inherited connective tissue disorders and investigating the role of the extracellular matrix in muscle.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{65}{Reelick, M.}{2011}
	{One step at a time. Disentangling the complexity of preventing falls in frail older persons.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{66}{Buur, P.F.}{2011}
	{Imaging in motion. Applications of multi-echo fMRI.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{67}{Schaefer, R.S.}{2011}
	{Measuring the mind’s ear: EEG of music imagery.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{68}{Xu, L.}{2011}
	{The non-preganglionic Edinger-Westphal nucleus: An integration center for energy balance and stress adaptation.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{69}{Schellekens, A.F.A.}{2011}
	{Gene-environment interaction and intermediate phenotypes in alcohol dependence.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{70}{Van Marle, H.J.F.}{2011}
	{The amygdala on alert: A neuroimaging investigation into amygdala function during acute stress and its aftermath.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{71}{De Laat, K.F.}{2011}
	{Motor performance in individuals with cerebral small vessel disease: An MRI study.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{2066/91394}
\dientry{72}{Mädebach, A.}{2011}
	{Lexical access in speaking: Studies on lexical selection and cascading activation.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{73}{Poelmans, G.J.V.}{2011}
	{Genes and protein networks for neurodevelopmental disorders.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{74}{Van Norden, A.G.W.}{2011}
	{Cognitive function in elderly individuals with cerebral small vessel disease. An MRI study.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{75}{Jansen, E.J.R.}{2011}
	{New insights into V-ATPase functioning: the role of its accessory subunit Ac45 and a novel brain-specific Ac45 paralog.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{76}{Haaxma, C.A.}{2011}
	{New perspectives on preclinical and early stage Parkinson's disease.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{77}{Haegens, S.}{2012}
	{On the functional role of oscillatory neuronal activity in the somatosensory system.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/92578}
\dientry{78}{van Barneveld, D.C.P.B.M.}{2012}
	{Integration of exteroceptive and interoceptive cues in spatial localization.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/91404}
\dientry{79}{Spies, P.E.}{2012}
	{The reflection of Alzheimer disease in CSF.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{80}{Helle, M.}{2012}
	{Artery-specific perfusion measurements in the cerebral vasculature by magnetic resonance imaging.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{81}{Egetemeir, J.}{2012}
	{Neural correlates of real-life joint action.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{82}{Janssen, L.}{2012}
	{Planning and execution of (bi)manual grasping.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{83}{Vermeer, S.}{2012}
	{Clinical and genetic characterisation of autosomal recessive cerebellar ataxias.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{84}{Vrins, S.}{2012}
	{Shaping object boundaries: Contextual effects in infants and adults.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{85}{Weber, K.M.}{2012}
	{The language learning brain: Evidence from second language and bilingual studies of syntactic processing.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{2066/92747}
\dientry{86}{Verhagen, L.}{2012}
	{How to grasp a ripe tomato.}
	{Utrecht University}{Utrecht}{the Netherlands}{}
\dientry{87}{Nonkes, L.J.P.}{2012}
	{Serotonin transporter gene variance causes individual differences in rat behaviour: For better and for worse.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{2066/93571}
\dientry{88}{Joosten-Weyn Banningh, L.W.A.}{2012}
	{Learning to live with Mild Cognitive Impairment: development and evaluation of a psychological intervention for patients with Mild Cognitive Impairment and their significant others.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{89}{Xiang, HD.}{2012}
	{The language networks of the brain.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{90}{Snijders, A.H.}{2012}
	{Tackling freezing of gait in Parkinson's disease.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{2066/93606}
\dientry{91}{Rouwette, T.P.H.}{2012}
	{Neuropathic pain and the brain - Differential involvement of corticotropin-releasing factor and urocortin 1 in acute and chronic pain processing.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{92}{Van de Meerendonk, N.}{2012}
	{States of indecision in the brain: Electrophysiological and hemodynamic reflections of monitoring in visual language perception.}
	{Radboud University Nijmegen}{Nijmegen}{the Netherlands}{}
\dientry{93}{Sterrenburg, A.}{2012}
	{The stress response of forebrain and midbrain regions: Neuropeptides, sex-specificity and epigenetics.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{94}{Uithol, S.}{2012}
	{Representing action and intention.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{95}{Van Dam, W.O.}{2012}
	{On the specificity and flexibility of embodied lexical-semantic representations.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{96}{Slats, D.}{2012}
	{ CSF biomarkers of Alzheimer’s disease: Serial sampling analysis and the study of circadian rhythmicity.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{97}{Van Nuenen, B.F.L.}{2012}
	{Cerebral reorganization in premotor parkinsonism.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{2066/100895}
\dientry{98}{van Schouwenburg, M.R.}{2012}
	{Fronto-striatal mechanisms of attentional control.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{99}{Azar, M.G.}{2012}
	{On the theory of reinforcement learning: Methods, convergence analysis and sample complexity.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{100}{Meeuwissen, E.B.}{2012}
	{Cortical oscillatory activity during memory formation.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/99202}
\dientry{101}{Arnold, J.F.}{2012}
	{When mood meets memory: Neural and behavioral perspectives on emotional memory in health and depression.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/105840}
\dientry{102}{Gons, R.A.R.}{2012}
	{Vascular risk factors in cerebral small vessel disease: A diffusion tensor imaging study.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{103}{Wingbermühle, E.}{2012}
	{Cognition and emotion in adults with Noonan syndrome: A neuropsychological perspective.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{104}{Walentowska, W.}{2012}
	{Facing emotional faces. The nature of automaticity of facial emotion processing studied with ERPs.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{105}{Hoogman, M.}{2012}
	{Imaging the effects of ADHD risk genes.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{106}{Tramper, J. J.}{2012}
	{Feedforward and feedback mechanisms in sensory motor control.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{107}{Van Eijndhoven, P.}{2012}
	{State and trait characteristics of early course major depressive disorder.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{108}{Visser, E.}{2012}
	{Leaves and forests: Low level sound processing and methods for the large-scale analysis of white matter structure in autism.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/105836}
\dientry{109}{Van Tooren-Hoogenboom, N.}{2012}
	{Neuronal communication in the synchronized brain. Investigating the functional role of visually-induced gamma band activity: Lessons from MEG.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{110}{Henckens, M.J.A.G.}{2012}
	{Imaging the stressed brain. Elucidating the time- and region-specific effects of stress hormones on brain function: A translational approach.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{111}{Van Kesteren, M.T.R.}{2012}
	{Schemas in the brain: Influences of prior knowledge on learning, memory, and education.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{112}{Brenders, P.}{2012}
	{Cross-language interactions in beginning second language learners.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{113}{Ter Horst, A.C.}{2012}
	{Modulating motor imagery. Contextual, spatial and kinaesthetic influences.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/105816}
\dientry{114}{Tesink, C.M.J.Y.}{2013}
	{Neurobiological insights into language comprehension in autism: Context matters.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/102332}
\dientry{115}{Böckler, A.}{2013}
	{Looking at the world together. How others’ attentional relations to jointly attended scenes shape cognitive processing.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{116}{Van Dongen, E.V.}{2013}
	{Sleeping to Remember. On the neural and behavioral mechanisms of sleep-dependent memory consolidation.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{117}{Volman, I.}{2013}
	{The neural and endocrine regulation of emotional actions.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{118}{Buchholz, V.}{2013}
	{Oscillatory activity in tactile remapping.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{119}{Van Deurzen, P.A.M.}{2013}
	{Information processing and depressive symptoms in healthy adolescents.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{120}{Whitmarsh, S.}{2013}
	{Nonreactivity and metacognition in mindfulness.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{121}{Vesper, C.}{2013}
	{Acting together: Mechanisms of intentional coordination.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/112295}
\dientry{122}{Lagro, J.}{2013}
	{Cardiovascular and cerebrovascular physiological measurements in clinical practice and prognostics in geriatric patients.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{the Netherlands}{}
\dientry{123}{Eskenazi, T.T.}{2013}
	{ You, us & them: From motor simulation to ascribed shared intentionality in social perception.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/112890}
\dientry{124}{Ondobaka, S.}{2013}
	{On the conceptual and perceptual processing of own and others’ behavior.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{125}{Overvelde, J.A.A.M.}{2013}
	{Which practice makes perfect? Experimental studies on the acquisition of movement sequences to identify the best learning condition in good and poor writers.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/116781}
\dientry{126}{Kalisvaart, J.P.}{2013}
	{Visual ambiguity in perception and action.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{127}{Kroes, M.}{2013}
	{Altering memories for emotional experiences.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{128}{Duijnhouwer, J.}{2013}
	{ Studies on the rotation problem in self-motion perception.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{129}{Nijhuis, E.H.J}{2013}
	{ Macroscopic networks in the human brain: Mapping connectivity in healthy and damaged brains.}
	{University of Twente}{Enschede}{The Netherlands}{}
\dientry{130}{Braakman, M. H.}{2013}
	{Posttraumatic stress disorder with secondary psychotic features. A diagnostic validity study among refugees in the Netherlands.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{131}{Zedlitz, A.M.E.E.}{2013}
	{Brittle brain power. Post-stroke fatigue, explorations into assessment and treatment.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{132}{Schoon, Y.}{2013}
	{From a gait and falls clinic visit towards self-management of falls in frail elderly.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{133}{Jansen, D.}{2013}
	{The role of nutrition in Alzheimer's disease - A study in transgenic mouse models for Alzheimer's disease and vascular disorders.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{134}{Kos, M.}{2013}
	{On the waves of language - Electrophysiological reflections on semantic and syntactic processing.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{135}{Severens, M.}{2013}
	{Towards clinical BCI applications: Assistive technology and gait rehabilitation.}
	{Radboud University Nijmegen, Nijmegen, Sint Maartenskliniek}{Nijmegen}{The Netherlands}{2066/112937}
\dientry{136}{Bergmann, H.}{2014}
	{Two is not always better than one: On the functional and neural (in)dependence of working memory and long-term memory.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{137}{Wronka, E.}{2013}
	{Searching for the biological basis of human mental abilitites. The relationship between attention and intelligence studied with P3.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{138}{Lüttjohann, A.K.}{2013}
	{The role of the cortico-thalamo-cortical system in absence epilepsy.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/115722}
\dientry{139}{Brazil, I.A.}{2013}
	{Change doesn’t come easy: Dynamics of adaptive behavior in psychopathy.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{140}{Zerbi, V.}{2013}
	{ Impact of nutrition on brain structure and function. A magnetic resonance imaging approach in Alzheimer mouse models.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{141}{Delnooz, C.C.S.}{2014}
	{Unravelling primary focal dystonia. A treatment update and new pathophysiological insights.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{2066/122939}
\dientry{142}{Bultena, S.S.}{2013}
	{Bilingual processing of cognates and language switches in sentence context.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{143}{Janssen, G.}{2014}
	{Diagnostic assessment of psychiatric patients: A contextual perspective on executive functioning.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{144}{Piai, V. Magalhães}{2014}
	{Choosing our words: Lexical competition and the involvement of attention in spoken word production.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{145}{Van  Ede, F.}{2014}
	{Preparing for perception. On the attentional modulation, perceptual relevance and physiology of oscillatory neural activity.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/124938}
\dientry{146}{Brandmeyer, A.}{2014}
	{Auditory perceptual learning via decoded EEG neurofeedback: a novel paradigm.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/125127}
\dientry{147}{Radke, S.}{2014}
	{Acting social: Neuroendocrine and clinical modulations of approach and decision behavior.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{148}{Simanova, I.}{2014}
	{In search of conceptual representations in the brain: towards mind-reading.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{149}{Kok, P.}{2014}
	{On the role of expectation in visual perception: A top-down view of early visual cortex.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{150}{Van Geldorp, B.}{2014}
	{The long and the short of memory: Neuropsychological studies on the interaction of working memory and long-term memory formation.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/126060}
\dientry{151}{Meyer, M.}{2014}
	{The developing brain in action - Individual and joint action processing.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{152}{Wester, A.}{2014}
	{Assessment of everyday memory in patients with alcohol-related cognitive disorders using the Rivermead Behavioural Memory Test.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{153}{Koenraadt, K.}{2014}
	{Shedding light on cortical control of movement.}
	{Radboud University Nijmegen, Nijmegen; Sint Maartenskliniek}{Nijmegen}{The Netherlands}{2066/125184}
\dientry{154}{Rutten-Jacobs, L.C.A.}{2014}
	{Long-term prognosis after stroke in young adults.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{155}{Herbert, M.K.}{2014}
	{ Facing uncertain diagnosis: the use of CSF biomarkers for the differential diagnosis of neurodegenerative diseases.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{156}{Llera Arenas, A.}{2014}
	{Adapting brain computer interfaces for non-stationary changes.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{157}{Smulders, K.}{2014}
	{ Cognitive control of gait and balance in patients with chronic stroke and Parkinson's disease.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{158}{Boyacioglu, R.}{2014}
	{On the application of ultra-fast fMRI and high resolution multiband fMRI at high static field strengths.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{159}{Kleinnijenhuis, M.}{2014}
	{Imaging fibres in the brain.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{160}{Geuze, J.}{2014}
	{Brain Computer Interfaces for Communication: Moving beyond the visual speller.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/127100}
\dientry{161}{Platonov, A.}{2014}
	{Mechanisms of binocular motion rivalry.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{}
\dientry{162}{Van der Schaaf, M.E.}{2014}
	{Dopaminergic modulation of reward and punishment learning.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{163}{Aerts, M.B.}{2014}
	{Improving diagnostic accuracy in parkinsonism.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{164}{Vlek, R.}{2014}
	{From Beat to BCI: A musical paradigm for, and the ethical aspects of Brain-Computer Interfacing.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/133272}
\dientry{165}{Massoudi, R.}{2014}
	{Interaction of task-related and acoustic signals in single neurons of monkey auditory cortex.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/128475}
\dientry{166}{Stolk, A.}{2014}
	{On the generation of shared symbols.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/129393}
\dientry{167}{Krause F.}{2014}
	{Numbers and magnitude in the brain: A sensorimotor grounding of numerical cognition.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/130434}
\dientry{168}{Munneke, M.A.M.}{2014}
	{Measuring and modulating the brain with non-invasive stimulation.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{169}{Von Borries, K.}{2014}
	{Carrots & Sticks - a neurobehavioral investigation of affective outcome processing in psychopathy.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{170}{Meel-van den Abeelen, A.S.S.}{2014}
	{In control. Methodological and clinical aspects of cerebral autoregulation and haemodynamics.}
	{Radboud University Nijmegen Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{171}{Leoné, F.T.M.}{2014}
	{Mapping sensorimotor space: Parieto-frontal contributions to goal-directed movements.}
	{Radboud University Nijmegen}{Nijmegen}{The Netherlands}{2066/131896}
\dientry{172}{Van Kessel, M.}{2014}
	{Nothing left? How to keep on the right track - Spatial and non-spatial attention processes in neglect after stroke.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/132138}
\dientry{173}{Vulto-van Silfhout, A. T.}{2014}
	{Detailed, standardized and systematic phenotyping for the interpretation of genetic variation.}
	{Radboud University Medical Centre}{Nijmegen}{The Netherlands}{}
\dientry{174}{Arnoldussen, D.}{2015}
	{Cortical topography of self-motion perception.}
	{Radboud University}{Nijmegen}{The Netherlands}{}
\dientry{175}{Meyer, M.C.}{2015}
	{“Inbetween Modalitiers combined EEG – fMR”.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/135188}
\dientry{176}{Bralten, J.}{2015}
	{Genetic factors and the brain in ADHD.}
	{Radboud university medical center}{Nijmegen}{The Netherlands}{}
\dientry{177}{Spaak, E.}{2015}
	{On the role of alpha oscillations in structuring neural information processing.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/141423}
\dientry{178}{Van der Doelen, R.}{2015}
	{Translational psychiatry; the twists and turns of early life stress and serotonin transporter gene variation.}
	{Radboud university medical center}{Nijmegen}{The Netherlands}{}
\dientry{179}{Lewis, C.}{2015}
	{The structure and function of intrinsic brain activity.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/139129}
\dientry{180}{Huang, Lili.}{2015}
	{The subiculum: a promising new target of deep brain stimulation in temporal lobe epilepsy. Investigation of closed-loop and open-loop high frequency stimulation of the subiculum in seizure and epilepsy models in rats.}
	{Radboud University}{Nijmegen}{The Netherlands}{}
\dientry{181}{Maaijwee, N.A.M.M}{2015}
	{Long-term neuropsychological and social consequences after stroke in young adults.}
	{Radboud university medical center}{Nijmegen}{The Netherlands}{}
\dientry{182}{Meijer, F.J.A.}{2015}
	{Clinical Application of Brain MRI in Parkinsonism: From Basic to Advanced Imaging.}
	{Radboud university medical center}{Nijmegen}{The Netherlands}{}
\dientry{183}{Van der Meij, R.}{2015}
	{On the identification, characterization and investigation of phase dependent coupling in neuronal networks.}
	{Radboud University}{Nijmegen}{The Netherlands}{}
\dientry{184}{Todorovic, A.}{2015}
	{Predictive adaptation in early auditory processing.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/141075}
\dientry{185}{Horschig, J.M.}{2015}
	{Flexible control and training of posterior alpha-band oscillations.}
	{Radboud University}{Nijmegen}{The Netherlands}{2066/140673}
\dientry{186}{Vogel, S.}{2015}
	{The runner-up: on the role of the mineralocorticoid receptor in human cognition.}
	{Radboud university medical center}{Nijmegen}{The Netherlands}{}
\dientry{187}{Clemens, I.A.H.}{2015}
	{Multi-sensory integration in location and self-motion perception.}
	{Radboud University}{Nijmegen}{The Netherlands}{}