### 2015-07-20 src/misc/s6_thanks.tex
Updated word of thanks.

\clearpage
\pagestyle{empty}

\chapter*{Acknowledgments}
\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgments}

Many people have contributed to my social and professional development over the course of my PhD. I would therefore like to extend my thanks to my friends, colleagues, and family for these wonderful six years.

\textbf{Pieter}, you are an amazing researcher and I'm proud to have been your student. I admire your sense of pragmatism and focus. Even though our (sometimes heated) discussions, I must admit that you have greatly influenced my writing style for the better. \textbf{Luc}, you greatly helped me iron out the detail in my experimental procedures, and provided crucial feedback when I was stuck analyzing our data. After receiving yet another batch of comments on one of my manuscripts, you were always there to provide support and to help me formulate a proper response. Many thanks to both of you!

In addition to \textbf{Pieter} and \textbf{Luc}, our collaborators \textbf{Mathieu}, \textbf{Maaike}, \textbf{Jan}, and \textbf{Paul} as well as the \textbf{technical support group} have played a crucial role in designing, setting up, running, and writing the experiments in this thesis. Thanks!

As with everything in life, writing this thesis was not always easy. Thanks to \textbf{Daphne} and my parents, \textbf{Marja} and \textbf{Wim}, for their unconditional support and confidence in me while I was writing this thesis.