### 2015-07-21 src/misc/s4_aboutme.tex
Updated about me.

\clearpage
%\pagestyle{empty}

\chapter*{Curriculum vitae}
\phantomsection
\addcontentsline{toc}{chapter}{Curriculum vitae}

Ivar Clemens was born in 1984 in Boxtel, the Netherlands. After obtaining a degree in Industrial Automation at the Koning Willem 1 College in 's-Hertogenbosch, he started a master's degree in Biology at the Radboud University in Nijmegen, the Netherlands. In 2008 Ivar did his first research internship on the \emph{integration of pre- and post-vergence target distance cues} with Pieter Medendorp at the Nijmegen Institute for Cognition and Information (NICI, later renamed to the Donders Centre for Cognition). Ivar then switched to the Donders Centre for Cognitive Neuro-imaging (DCCN) for his second research internship. This time he worked with  Ole Jensen and Saskia Haegens on \emph{homeostatic plasticity in the somatosensory cortex}. He obtained his master's degree in 2009 and started his PhD research under the supervision of Pieter Medendorp and Luc Selen at the Donders Centre for Cognition. During his PhD he worked on the integration of sensory signals in spatial orientation and self-motion perception. In January 2015 he joined the group dynamics lab at the Champalimaud Centre for the Unknown in Lisbon, Portugal.